Supplementary Information

### 2016-09-01 supplementary/supplementary.tex
new pdf

\usepackage{subfigure} 
\usepackage{floatrow}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{url}

@@

\begin{table}[h]\caption{AUC-PR upon addition of a new task - HIV, in the 10\% setting. Note that STL and the homolog method build independent models so their AUC numbers are the same as in Table \ref{resultsTable}. Among the multitask learning methods, BSL-MTL's performance is not hurt by the addition of an unrelated task such as HIV.}
\label{hivresults}
\begin{small}
\begin{center}
%\def\arraystretch{1.2}
\begin{tabular}{c|cccc}
\toprule
& \multicolumn{4}{c}{10 \% training } \\
& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} & \textit{HIV} \\ \midrule
Homolog & 0.230$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 & 0.170 $\pm$.01 \\
STL  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02  & 0.208 $\pm$.01 \\ 
Sparse+LR & 0.135 $\pm$.06 & 0.724 $\pm$.05 & 0.280 $\pm$.01 & 0.205 $\pm$.03  \\ 
BSL-MTL & \textbf{0.241} $\pm$.01 & \textbf{0.793} $\pm$.06 & \textbf{0.459} $\pm$.01 & \textbf{0.354} $\pm$.01  \\ \bottomrule
\end{tabular}
\end{center}
\end{small}
\end{table}

@@

Note that our features do not directly correspond to a unique amino-acid $k$-mer (see Section \ref{sec:datasets}):
the virus feature $f_v$ will map to several amino-acid sequences (for instance KKCC, KRCC, RRCC etc all map to a single feature due to the molecular similarity between the amino acids K and R being both positively charged). Given the set of top virus features we can obtain the corresponding set of
amino-acid $k$-mers, say $AA_v$, by reversing the feature-generation step. However most of the possible $k$-mers do not appear in the
training data (ex: out of the 160,000 (=$20^4$) possible 4-mers $\approx$24,000 appear). Let $AA_{tr}$
be the set of amino-acid k-mers that appear in the training data. Then, the intersection $I_v = AA_v \cap AA_{tr}$ gives us the important amino-acid $k$-mers from virus proteins w.r.t interaction prediction.
To summarize $I_v$, we use a popular tool Seq2Logo \citep{seq2logo} to generate a sequence motif.
The logos for the two-, three-, four-mers from $I_v$ are generated independently. Since we only
want to summarize, we use the Shannon logo type (which does not consider any background amino-acid distribution)
with the following settings: clustering-method=None, weight-on-prior=1 (pseudo-counts do not make sense in our
analysis). Figure \ref{fig:virus_motifs} shows the motif that is common across viruses. %, while the other figures highlight those motifs important in a specific species.
We observe that the shared tri-mer motif for virus proteins in Figure \ref{fig:virus_motifs} is dominated by
hydrophilic amino acids (T, K, R, D, E).

@@

 \noindent\emph{Phosphorylation sites}: 
 We found the frequent occurrence of S and T and sometimes Y in the motifs striking and suspected this may be
 related to the amino acids being frequent targets of phosphorylation.
 Phosphorylated sites often serve as PPI sites, and databases such as Phosphosite \cite{phosphosite} are
  repositories for known sites in human proteins. Since these are sites in human proteins, we searched for the patterns
	from the 4-mer motif in Fig \ref{fig:human_motifs} and found several to be flanking known phosphorylation sites in human proteins:
	\texttt{LLLs}, \texttt{LLLt}, \texttt{ILLs}, \texttt{PPPs}, \texttt{PIPs}, \texttt{PIPt}, \texttt{LIPs}, \texttt{PLLt} (lower-case indicates the putative phosphorylation site). This observation also supports the notion that the motifs predictive
	of interaction are biologically significant.

\noindent\emph{Evidence in IEDB}\footnote{\url{www.iedb.org}}:
We found experimental evidence for the significance for the virus motifs in the Immune Epitope\footnote{An epitope is a very short sequence from the virus that binds to human antibodies} database (IEDB) \cite{iedb}.
The pattern \texttt{IVGG} from the \textit{Hepatitis-C} motif in Fig. \ref{fig:hepc_motifs} is found in 53 epitopes.
From the \textit{Ebola} motifs in Fig \ref{fig:ebola_motifs}, we find that \texttt{TLAT} is part of six different epitopes, \texttt{SLTT} appears in three epitopes. \texttt{PLIK}, \texttt{SLLL} from the \textit{Influenza} motif are also found in many epitopes.
Finding that the virus k-mer patterns predicted by our method are recognized by human antibodies
is a further validation of its performance.
Further, using higher dimensional $k$-mers (where $k$=7, 8, 9) as
features in our model will give motifs from which complete epitopes can be derived. % which can then be verified by biochemical peptide interaction assays.
Our model thus has applications in epitope prediction as well, where conventional methods consist of scanning all possible $k$-mers from protein sequences to identify likely epitopes.

@@

\begin{figure}[h]
\begin{floatrow}
\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.25, trim = 0 0 0 0]{top10ebola_3mers_seq2logo.eps}
}{%
\caption{Sequence motifs specific to \textit{Ebola} proteins that are also important to interactions.}%
\label{fig:hepc_motifs}
}

\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.25, trim = 0 0 0 0]{top10flavi_3mers_seq2logo.eps}
}{%
\caption{Sequence motifs specific to \textit{Hepatitic-C} proteins that are also important to interactions.}%
\label{fig:hepc_motifs}
}
\end{floatrow}
\end{figure}

\begin{figure}[h]
\begin{floatrow}
\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.25, trim = 0 0 0 0]{top10flu_3mers_seq2logo.eps}
}{%
\caption{Sequence motifs specific to \textit{Influenza} proteins that are also important to interactions.}%
\label{fig:flu_motifs}
}
\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.25, trim = 0 0 0 0]{top20human_3mers_seq2logo.eps}
}{%
\caption{Sequence motif constructed from the top tri-mer features of human proteins}%
\label{fig:human_motifs}
}
\end{floatrow}
\end{figure}

>>>>

\usepackage{floatrow}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{subfigure} 
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{url}

\newcommand{\argmin}{\operatornamewithlimits{argmin}}

@@

\begin{table}[h]\caption{AUC-PR upon addition of a new task - HIV, in the 10\% setting. Note that STL and the homolog method build independent models so their AUC numbers are the same as in Table 2. Among the multitask learning methods, BSL-MTL's performance is not hurt by the addition of an unrelated task such as HIV.}
\label{hivresults}
\begin{small}
\begin{center}
%\def\arraystretch{1.2}
\begin{tabular}{c|cccc}
\toprule
& \multicolumn{4}{c}{10 \% training } \\
& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} & \textit{HIV} \\ \midrule
Homolog & 0.230$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 & 0.170 $\pm$.01 \\
STL  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02  & 0.208 $\pm$.01 \\ 
Sparse+LR & 0.135 $\pm$.06 & 0.724 $\pm$.05 & 0.280 $\pm$.01 & 0.205 $\pm$.03  \\ 
BSL-MTL & \textbf{0.241} $\pm$.01 & \textbf{0.793} $\pm$.06 & \textbf{0.459} $\pm$.01 & \textbf{0.354} $\pm$.01  \\ \bottomrule
\end{tabular}
\end{center}
\end{small}
\end{table}

@@

Note that our features do not directly correspond to a unique amino-acid $k$-mer (see Section 4.2):
the virus feature $f_v$ will map to several amino-acid sequences (for instance KKCC, KRCC, RRCC etc all map to a single feature due to the molecular similarity between the amino acids K and R being both positively charged). Given the set of top virus features we can obtain the corresponding set of
amino-acid $k$-mers, say $AA_v$, by reversing the feature-generation step. However most of the possible $k$-mers do not appear in the
training data (ex: out of the 160,000 (=$20^4$) possible 4-mers $\approx$24,000 appear). Let $AA_{tr}$
be the set of amino-acid k-mers that appear in the training data. Then, the intersection $I_v = AA_v \cap AA_{tr}$ gives us the important amino-acid $k$-mers from virus proteins w.r.t interaction prediction.
To summarize $I_v$, we use a popular tool Seq2Logo to generate a sequence motif.
The logos for the two-, three-, four-mers from $I_v$ are generated independently. Since we only
want to summarize, we use the Shannon logo type (which does not consider any background amino-acid distribution)
with the following settings: clustering-method=None, weight-on-prior=1 (pseudo-counts do not make sense in our
analysis). Figure \ref{fig:virus_motifs} shows the motif that is common across viruses. %, while the other figures highlight those motifs important in a specific species.
We observe that the shared tri-mer motif for virus proteins in Figure \ref{fig:virus_motifs} is dominated by
hydrophilic amino acids (T, K, R, D, E).

@@

 \noindent\emph{Phosphorylation sites}: 
 We found the frequent occurrence of S and T and sometimes Y in the motifs striking and suspected this may be
 related to the amino acids being frequent targets of phosphorylation.
 Phosphorylated sites often serve as PPI sites, and databases such as Phosphosite are
  repositories for known sites in human proteins. Since these are sites in human proteins, we searched for the patterns
	from the 4-mer motif in Fig \ref{fig:human_motifs} and found several to be flanking known phosphorylation sites in human proteins:
	\texttt{LLLs}, \texttt{LLLt}, \texttt{ILLs}, \texttt{PPPs}, \texttt{PIPs}, \texttt{PIPt}, \texttt{LIPs}, \texttt{PLLt} (lower-case indicates the putative phosphorylation site). This observation also supports the notion that the motifs predictive
	of interaction are biologically significant.

\noindent\emph{Evidence in IEDB}\footnote{\url{www.iedb.org}}:
We found experimental evidence for the significance for the virus motifs in the Immune Epitope\footnote{An epitope is a very short sequence from the virus that binds to human antibodies} database (IEDB).
The pattern \texttt{IVGG} from the \textit{Hepatitis-C} motif in Fig. \ref{fig:hepc_motifs} is found in 53 epitopes.
From the \textit{Ebola} motifs in Fig \ref{fig:ebola_motifs}, we find that \texttt{TLAT} is part of six different epitopes, \texttt{SLTT} appears in three epitopes. \texttt{PLIK}, \texttt{SLLL} from the \textit{Influenza} motif are also found in many epitopes.
Finding that the virus k-mer patterns predicted by our method are recognized by human antibodies
is a further validation of its performance.
Further, using higher dimensional $k$-mers (where $k$=7, 8, 9) as
features in our model will give motifs from which complete epitopes can be derived. % which can then be verified by biochemical peptide interaction assays.
Our model thus has applications in epitope prediction as well, where conventional methods consist of scanning all possible $k$-mers from protein sequences to identify likely epitopes.

@@

\begin{figure}[h]
\begin{floatrow}
\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.25, trim = 0 0 0 0]{top10ebola_3mers_seq2logo.pdf}
}{%
\caption{Sequence motifs specific to \textit{Ebola} proteins that are also important to interactions.}%
\label{fig:hepc_motifs}
}

\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.25, trim = 0 0 0 0]{top10flavi_3mers_seq2logo.pdf}
}{%
\caption{Sequence motifs specific to \textit{Hepatitic-C} proteins that are also important to interactions.}%
\label{fig:hepc_motifs}
}
\end{floatrow}
\end{figure}

\begin{figure}[h]
\begin{floatrow}
\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.25, trim = 0 0 0 0]{top10flu_3mers_seq2logo.pdf}
}{%
\caption{Sequence motifs specific to \textit{Influenza} proteins that are also important to interactions.}%
\label{fig:flu_motifs}
}
\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.25, trim = 0 0 0 0]{top10human_3mers_seq2logo.pdf}
}{%
\caption{Sequence motif constructed from the top tri-mer features of human proteins}%
\label{fig:human_motifs}
}
\end{floatrow}
\end{figure}
